The impact of forest management on the temperature sensitivity of organic matter turnover

### 2024-11-15 Manuscript_draft/Draft.Rmd
discussion

## Forest management effect on activation energy and sensitivity to temperature
Generally, our results confirmed the hypothesis 1, since slash addition increased the respiration rates compared to the treatments where the slash was removed. Addition of slash created an artificial litter layer that buffered the effects over soil microclimatic conditions, and provided new habitat and food availability for soil organisms (ref). These results indicate that increases in heterotrophic respiration due to due to an increment in organic matter availability by slash addition compensated for the decreases in autotrophic respiration and root-associated microbiota due to tree logging and soil disturbance (Borkhuu et al. 2015). Moreover, the strength of the response was also dependent on the forest, being more evident in Spain and France where the tree density was much higher than Romania. In Romania such a low density of trees together with a dense understory of grasses in all treatments might explain the lack of response and indicate high degree of resilience to perturbations.

@@

## Role of edaphic, ecological and microbiological variables in temperature sensitivity
The Q10 values obtained in our study were comparable to the Q10 values measured in the field by other studies (ref). Xu et al 2015 observed the largest relative frequency of Q10 values across forest in China was within the range of 2.0–3.0. Fierer et al. (2006) observed Q10 values ranging from 2.2 to 4.6 (average of 3.0) at continental-scale in an incubation experiment. Still, we observed a large variation of Q10 depending on the forest, being the highest in Romania. This is in agreement with the known higher temperature sensitivity in colder regions (Koven et al 2017). The reason could be related to longer time of frozen soils that could reduce the diffusion of substrates and extracellular enzymes within the soil (Davinson and Jansen 2006), giving higher values of Q10 not related to recalcitrance.

Our analyses revealed the three variables that explained most of the variation of Q10 were related to i) the quality of substrate (i.e., nitrogen content as limiting nutrient in forest), ii) the microorganisms, in particular, fungal to bateria ratio and iii) the primary productivity (that is, the number of trees).

>>>>

## Forest management effect on activation energy and sensitivity to temperature
Generally, our results confirmed the hypothesis 1, since slash addition increased the respiration rates compared to the treatments where the slash was removed. Addition of slash created an artificial litter layer that buffered the effects over soil microclimatic conditions, and provided new habitat and food availability for soil organisms (ref). These results indicate that increases in heterotrophic respiration due to an increment in organic matter availability by slash addition compensated for the decreases in autotrophic respiration and root-associated microbiota due to tree logging and soil disturbance (Borkhuu et al. 2015). Moreover, the strength of the response was also dependent on the forest, being more evident in Spain and France where the tree density was much higher than Romania. In Romania such a low density of trees together with a dense understory of grasses in all treatments might explain the lack of response and indicate high degree of resilience to perturbations.

@@

## Role of edaphic, ecological and microbiological variables in temperature sensitivity
The Q10 values obtained in our study were comparable to the Q10 values measured by other studies. Xu et al 2015 observed the largest relative frequency of Q10 values across forest in China was within the range of 2.0–3.0. Fierer et al. (2006) observed Q10 values ranging from 2.2 to 4.6 (average of 3.0) at continental-scale in an incubation experiment. Still, we observed a large variation of Q10 depending on the forest, being the highest in Romania. This is in agreement with the known higher temperature sensitivity in colder regions (Koven et al 2017). Longer time of frozen soils in Romania could have reduced the diffusion of substrates and extracellular enzymes within the soil (Davinson and Jansen 2006), giving higher values of Q10 not related to recalcitrance.

Our analyses revealed the three variables that explained most of the variation of Q10 were related to i) the quality/quantity of substrate (i.e., nitrogen content), ii) the microorganisms, in particular, fungal to bacteria ratio and iii) the soil temperature. Nitrogen fertilization experiments have shown that nitrogen availability decreases the Q10 (Mo et al 2008; Zhang et al 2019), probably related to changes in microbial carbon use efficiency (Manzoni et al 2012). In our experiment, it seems that addition of slash have stimulated decomposition of labile C which increased N availability (Ghosh et al., 2019; Shahbaz et al.,2017), and thus decreased the temperature sensitivity. Our results indicate that SOM decomposition not only depends on the substrate biochemistry but also on the ability of the microbial community to decompose the available substrate. This is in agreement with a recent global-scale study stating that soil microorganism are the main players governing the responses of soil heterotrophic respiration to temperature, and secondly substrate quantity and biochemical recalcitrance (Sáez-Sandino et al 2023).

We did not measure chemical recalcitrance per se, but our results suggest that microbial communities might have a key role in determining Q10 in forest ecosystems. Lower Q10 values associated to the addition of fresh plant material as slash was related to higher fungal to bateria rates. Although both fungi and bacteria have the capacity to decompose litter components (Stursova et al. 2012), fungi has a major role in the decomposition of plant biomass due to their mycelial growth, the ability to translocate nutrients and the possession of a strong enzymatic capacity (Schneider et al. 2012, Tlaskal et al 2016). In contrast, bacteria seems to preferentially gain carbon from live or death fungal biomass and increase at later stages of litter decomposition (Tlaskal et al 2016, Lindahl et al 2010). Since ectomycorrhizal fungal mycelium comprise most of the fungal biomass in forests (Wallander et al 2001), a strong reduction in ectomycorrhizal fungal biomass is expected in all the tree removal disturbance treatments, thus decreasing the fungal to bacteria ratio. As mentioned, this decrease in ectomycorrhizal fungal biomass might be compensated by a rapid growth of saprotrophic fungal species where slash is added. Community data would give more insights about that. Surprinsingly, hydrolitic enzymes were not the main variables affecting Q10. This effect might be masked by the positive correlation of fungal and bacteria biomass with enzymes. Still, most enzymes correlated negatively with Q10. The exception was beta-xylosidase, that correlated positively with Q10 and explained a high proportion of variance. The only microbial group that strongly correlated with beta-xylosidase was the actinobacteria and problably the relationship is determined by the exceptionally high activity of these enzyme in Romania, compared with other sites. This result is consistent with previous studies showing that members of the phylum Actinobacteria are oligotrophic micoorganisms with the ability to compete for resources in a resource-limited environment (Trivedi et al 2013; Bao et al 2021).
